---
title: "Creating_New_Variables"
author: "I-Ting Yu"
date: "August 20, 2016"
output: html_document
---

## Why creating new variables?
* Raw data wont have values that you're looking for
* Tranferm data to what you desire
* Add some varaibles to the data frame that you're working with

Common variables:
- Missing indicators  
- "Cutting up" quantitative varaibles (factor version) 
- Applying transformations 
```{r}
if(!file.exists("./data/restaurants.csv")) {
    fileUrl <- "https://data.baltimorecity.gov/api/views/k5ry-ef3g/rows.csv?accessType=DOWNLOAD"
    download.file(fileUrl, destfile = "./data/restaurants.csv", method = "curl")
}

rest.data <- read.csv("./data/restaurants.csv")
```

### Create sequences

Often used to index different operations that you're gonna to do on data
```{r}
s1 <- seq(1, 10, by = 2)
s1
s2 <- seq(1, 10, length = 3)
s2

# Create index in order to loop through the data
x <- c(6, 8, 1, 28,13)
seq(along = x)
```

### Subsetting varaibles
```{r}
rest.data$nearMe <- rest.data$neighborhood %in% c("Roland Park", "Homeland")
head(rest.data)
table(rest.data$nearMe)
```
Make it a lot earier when we only want to deal the data in our neighborhood("Roland Park", "Homeland")  
i.e. no longer have to use : rest.data$neighborhood %in% c("Roland Park", "Homeland")

### Create binary variables
```{r}
rest.data$zipWrong <- ifelse(rest.data$zipCode < 0, yes = TRUE, no = FALSE)
table(rest.data$zipWrong)
```

### Create categorical variable

```{r}
rest.data$zipGroups <- cut(rest.data$zipCode, breaks = quantile(rest.data$zipCode))

table(rest.data$zipGroups)
table(rest.data$zipGroups, rest.data$zipCode)

# easier cutting
library(Hmisc)
# g: number of quantitle groups
rest.data$zipGroups <- cut2(rest.data$zipCode, g = 4)
table(rest.data$zipGroup)
```


### Create factor variable
```{r}
rest.data$zcf <- factor(rest.data$zipCode)
rest.data$zcf[1:10]
```

#### Use the mutate function
```{r}
library(Hmisc)
library(plyr)
rest.data.2 <- mutate(rest.data, zipGroups = cut2(zipCode, g = 4))
head(rest.data.2)
```

Common transformatons:  
* round(x , digit = n)  
* signif(x, digit = n)  
* log(x)  
* log2(x)  
* log10(x)  
* exp(x)

