---
title: "Data_Table"
author: "I-Ting Yu"
date: "August 16, 2016"
output: html_document
---
  
```{r}
DF <- data.frame(x = rnorm(9), y = rep(c("a", "b", "c"), each = 3), z = rnorm(9))
head(DF,3)
class(DF)

```

Inherets from data.frame
```{r}
library(data.table)
DT <- data.table(x = rnorm(9), y = rep(c("a", "b", "c"), each = 3), z = rnorm(9))
head(DT,3)
class(DT)
```

```{r}
# See all data tables in memory
tables()

# Subsetting 
DT[2, ]
DT[DT$y == "a", ]
DT[c(2,3)]
DT[c(2,3), ]

# NOTICE Subsetting columns??! 
DT[, c(2,3)]
```

### Calcualting values for variables with expressions
The argument you pass after the comma is called an expression(e.g. DT[, mean(x)] )
```{r}
# Same results
print(DT)
mean(DT$x)
DT[, mean(x)]

DT[, list(mean(x), sum(z))]

table(DT$y)
DT[, table(y)]
```

### Adding new columns
It doesnt create a new copy like data.frame. More memory effient, especially dealing with large data set
```{r}
DT[, w := z^2]
print(DT)
```

Careful!!  
Because the new copy hasn't been made. The changes reflect on DT2 as well.
```{r}
DT2 <- DT
head(DT2, 3)

DT[, y := 2]
head(DT, 3)
head(DT2, 3)

```

### Manipulate coulums

```{r}
DT[, m := { tmp <- (x+z); tmp + 1}]
head(DT, 3)

# plyr like operation
DT[, a := x > 0]
head(DT, 3)

DT[, b := mean(x + w), by = a]
head(DT, 3)
```

### Special Variable
".N" counts the number of observation of a unique level by column
```{r}
set.seed(123)
DT <- data.table(
      x = sample(letters[1:5], size = 100, replace = TRUE),
      y = sample(LETTERS[1:5], size = 100, replace = TRUE))
head(DT, 10)

DT[, .N, by = x]
DT[, .N, by = y]

```

### Keys

```{r}
DT <- data.table(
          x = rep(c("c", "b", "a"), each = 5),
          y = rnorm(15))

setkey(DT, x)

# Subset data by key
DT['a']

```


### Joins
```{r}
DT1 <- data.table(x = c("a", "a", "b", "dt1"), y = 1:4)
DT1
DT2 <- data.table(x = c("b", "a", "dt2"), z = 5:7)
DT2

setkey(DT1, x)
setkey(DT2, x)

# Remember to set keys from 2 tables to be joined
merge(DT1, DT2)

```

### Fast Reading 
```{r}
# Using fread, 10x faster 
big.df <- data.frame(x = rnorm(1E5), y = rnorm(1E5))
tmp.file <- tempfile()
write.table(big.df, file = tmp.file, row.names = FALSE, col.names = TRUE, sep = "\t", quote = FALSE)
system.time(fread(tmp.file))

# Reguler way
system.time(read.table(tmp.file, header = TRUE, sep = "\t"))

```




