---
title: "get-n-clean-data-quiz1.rmd"
author: "I-Ting Yu"
date: "August 16, 2016"
output: html_document
---

## Question 1
The American Community Survey distributes downloadable data about United States communities. Download the 2006 microdata survey about housing for the state of Idaho using download.file() from here:

https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06hid.csv

and load the data into R. The code book, describing the variable names is here:

https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FPUMSDataDict06.pdf

How many properties are worth $1,000,000 or more?

```{r}
if (!file.exists("data")) {
    dir.create("data")
}

fileUrl <- "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06hid.csv"
download.file(fileUrl, destfile = "./data/06hid.csv", method = "curl")
dataDownloaded <- date()

data <- read.csv("./data/06hid.csv")
head(data, 3)

idx.col <- which(colnames(data)=="VAL")
data.val <- data[data$VAL == 24, idx.col]
sum(!is.na(data.val))

```

## Question 2 (skipped)  

## Question 3
Download the Excel spreadsheet on Natural Gas Aquisition Program here:  

https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FDATA.gov_NGAP.xlsx

Read rows 18-23 and columns 7-15 into R and assign the result to a variable called: dat  

What is the value of: sum(dat$Zip * dat$Ext, na.rm = T)

```{r}
library(xlsx)
fileUrl <- "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FDATA.gov_NGAP.xlsx"

download.file(fileUrl, destfile = "./data/gas.xlsx", method = "curl")
dateDownloaded <- date()

# Troubleshoot: install.packages("rjava")
# Troubleshoot: install.packages("xlsxjars")
dat <- read.xlsx("./data/gas.xlsx", sheetIndex = 1, rowIndex = 18:23, colIndex = 7:15 )

sum(dat$Zip * dat$Ext, na.rm = TRUE)
```

## Question 4
Read the XML data on Baltimore restaurants from here:

https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Frestaurants.xml

How many restaurants have zipcode 21231?

##### Becuase xmlTreeParse does not support "https"
Option 1: change "https" to "http"
```{r}
library(XML)
fileUrl <- "http://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Frestaurants.xml"

download.file(fileUrl, destfile = "./data/Bal_res", method = "curl")
dateDownloaded <- date()

doc <- xmlTreeParse(fileUrl, useInternal = TRUE)

```

Option 2: use Rcurl
```{r}
library(RCurl)
fileUrl <- "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Frestaurants.xml"
doc.byRCurl <- getURL(fileUrl)
doc.byRCurl <- xmlTreeParse(doc.byRCurl, useInternal = TRUE)

```

```{r}
# Wrapper for the entire document 
rootNode <- xmlRoot(doc)
xmlName(rootNode)

zipcodes <- xpathSApply(rootNode, "//zipcode", xmlValue)

# The previous returns a chararter vector 
class(xpathSApply(rootNode, "//zipcode", xmlValue))

sum(zipcodes == 21231)

```

## Question 5
The American Community Survey distributes downloadable data about United States communities. Download the 2006 microdata survey about housing for the state of Idaho using download.file() from here:

https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06pid.csv

using the fread() command load the data into an R object: DT

Which of the following is the fastest way to calculate the average value of the variable pwgtp15 broken down by sex using the data.table package?

* DT[,mean(pwgtp15),by=SEX]
* tapply(DT$pwgtp15,DT$SEX,mean)
* mean(DT$pwgtp15,by=DT$SEX)
* rowMeans(DT)[DT$SEX==1]; rowMeans(DT)[DT$SEX==2]
* sapply(split(DT$pwgtp15,DT$SEX),mean)
* mean(DT[DT$SEX==1,]$pwgtp15); mean(DT[DT$SEX==2,]$pwgtp15)

```{r}
library(data.table)

fileUrl <- "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06pid.csv"

download.file(fileUrl, destfile = "./data/fss06pid.csv", method="curl")

dateDownloaded <- date()
```

```{r}
DT <- fread(input = "./data/fss06pid.csv")

system.time(DT[, mean(pwgtp15), by = SEX])

system.time(tapply(DT$pwgtp15,DT$SEX, mean))
  
system.time(mean(DT$pwgtp15,by = DT$SEX))

system.time(sapply(split(DT$pwgtp15,DT$SEX), mean))
  
system.time(mean(DT[DT$SEX == 1, ]$pwgtp15))

system.time(mean(DT[DT$SEX == 2, ]$pwgtp15))

```




