---
title: "get-n-clean-data-quiz-2"
author: "I-Ting Yu"
date: "August 19, 2016"
output: html_document
---

#### Question 1 
Register an application with the Github API here https://github.com/settings/applications  
Access the API to get information on your instructors repositories  https://api.github.com/users/jtleek/repos      
Use this data to find the time that the datasharing repo was created.  What time was it created? This tutorial may be useful  https://github.com/hadley/httr/blob/master/demo/oauth2-github.r  

```{r}
library(httr)
library(httpuv)
library(jsonlite)
```

```{r}
client.id <- "bc1ef97b9fe0e5a2b492"
client.secret <- "d05449f9535e64091820a7670131328e16523682"

```

```{r}
# 1. OAuth Settings

#client.id <- "yourClientID"
#client.secret <- "yourClientSecret"
oauth_endpoints("github")
myapp <- oauth_app(appname = "github", key = client.id, secret = client.secret)
print(myapp)
print(class(myapp))

# 2. Get OAuth credentials
github.token <- oauth2.0_token(endpoint = oauth_endpoints("github"), app = myapp)
print(github.token)
print(class(github.token))

# 3. Access API
g.request.token <- config(token = github.token)
print(g.request.token)
print(class(g.request.token))

json.response <- GET("https://api.github.com/users/jtleek/repos", config = g.request.token)
print(json.response)
print(class(json.response))
stop_for_status(json.response)

# 4. Extract out the content from the json.response
content.list = content(x = json.response)
print(class(content.list))

# 5. convert the list to json 
# Basically the list has json-formatted content, but the formatting make it hard to read. As a result, we use "toJSON" to parse the list and use "fromJSON" to convert the json content into a data.frame  
my.data = fromJSON(toJSON(content.list))
print(class(my.data))

# 6. Result 
my.data[my.data$full_name == "jtleek/datasharing", ]$created_at
```

#### Question 2 
The sqldf package allows for execution of SQL commands on R data frames. We will use the sqldf package to practice the queries we might send with the dbSendQuery command in RMySQL. Download the American Community Survey data and load it into an R object called

acs  

https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06pid.csv

Which of the following commands will select only the data for the probability weights that are formatted like pwgtp1, pwgtp2, pwgtp3, etc. for the people with ages less than 50?

```{r}
library(sqldf)
library(RMySQL)
fileUrl <- "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06pid.csv"

download.file(url = fileUrl, destfile = "./data/acs.csv")

acs <- read.csv("./data/acs.csv")

head(acs)

query.res.df <- sqldf("SELECT pwgtp1 FROM acs WHERE AGEP < 50", drv = "SQLite")

head(query.res.df)
```

#### Question 3
Using the same data frame you created in the previous problem, what is the equivalent function to unique(acs$AGEP)

```{r}
sqldf("SELECT distinct AGEP FROM acs", drv = "SQLite")
```

#### Question 4 
How many characters are in the 10th, 20th, 30th and 100th lines of HTML from this page:

[link](http://biostat.jhsph.edu/~jleek/contact.html) http://biostat.jhsph.edu/~jleek/contact.html

(Hint: the nchar() function in R may be helpful)

```{r}
conn <- url("http://biostat.jhsph.edu/~jleek/contact.html")

# each line is a character vector
html.lines <- readLines(con = conn)

# traversing over a set of data like a list or "vector"
sapply(html.lines[c(10, 20, 30, 100)], nchar)
```

#### Question 5 
Read this data set into R and report the sum of the numbers in the fourth column.  

https://d396qusza40orc.cloudfront.net/getdata%2Fwksst8110.for  

Original source of the data:  
http://www.cpc.ncep.noaa.gov/data/indices/wksst8110.for  

(Hint this is a fixed width file format)

```{r}
fileUrl <- "http://www.cpc.ncep.noaa.gov/data/indices/wksst8110.for"

download.file(fileUrl, destfile="./sst8110.for")

# The -1 in the widths argument means there is a one-character column that should be ignored, the -5 means a five-character column that should be ignored.
the.widths <- c(-1, 9, -5, 4, 4, -5, 4, 4, -5, 4, 4, -5, 4, 4)
my.data <- read.fwf(file = "./sst8110.for",
                    skip = 4,
                    widths = the.widths)
head(my.data)

sum(my.data[, 4])
```



